---
title: "QAPP_Plots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  {.tabset}

plots and tables for the 2022 lci qapp

## Parameter History


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
#merging lake id with the DOH ID
library(tidyverse)
rm(list=setdiff(ls(), c("newdata")))
# setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/Current")
# source("new_database/Reading.LMAS.Data.R")
# setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/QAPP_Plots")

lci<-newdata %>% filter(DATA_PROVIDER=="LCI",
                        SAMPLE_TYPE=="WATER COLUMN") %>% 
  mutate(year=substr(SAMPLE_DATE,1,4),
         class=ifelse(grepl("A",LOCATION_WATERBODY_CLASSIFICATION),"A","B"),
         info=INFORMATION_TYPE,
         param=CHARACTERISTIC_NAME) %>% 
  group_by(param,year) %>% 
  mutate(core=ifelse(length(unique(SAMPLE_DATE))>5,"yes","no")) %>% 
  ungroup() %>% 
  #remove miscategoriezed depth profile data
  mutate(info=ifelse(info=='DP'&RSLT_RESULT_TYPE=='LAB',NA,info)) %>% 
  filter(!is.na(info)) %>% 
  select(year,param,info,class,core,RSLT_RESULT_SAMPLE_FRACTION) %>% distinct() %>% arrange(year,param)

#preserve list of parameters that aren't core (meaning sampled < 5x/year)
lci_not_core_params<-lci %>% filter(core=="no")
lci<-lci %>% filter(core=="yes") %>% select(-core)

#add in param groupings and remove those without groupings because not important
param_groups<-read.csv("param.groupings.csv",na.strings=c("","NA"))
param_groups<-param_groups %>% rename(param=CHARACTERISTIC_NAME) %>% 
  select(category,param,RSLT_RESULT_SAMPLE_FRACTION,simple) %>% distinct()
lci<-merge(lci,param_groups,by=c('param','RSLT_RESULT_SAMPLE_FRACTION'),all.x=TRUE)
lci<-lci %>% filter(!is.na(category))

#add in 2022 proposed sampling
new<-read.csv("param.groupings.csv",na.strings=c("","NA"))
new<-new %>% 
  rename(param=CHARACTERISTIC_NAME) %>% 
  mutate(year="2022") %>%
  pivot_longer(    cols = c('DP','OW','BS'),
    names_to = "info",
    values_to = "remove",
    values_drop_na = TRUE) %>% 
  select(-remove) %>% distinct() %>% 
  pivot_longer(    cols = c('A','B'),
    names_to = "class",
    values_to = "remove",
    values_drop_na = TRUE) %>% 
  select(-remove) %>% distinct()
lci<-merge(lci,new,all=TRUE)

```

plot of parameter history

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
library(ggplot2)

salts<-lci %>% filter(category=="Salt_Parameters") %>% 
  mutate(layer_class=paste(info,class,sep="_"))

ggplot(salts,
       aes(x = simple,y = year)) + 
  geom_point(aes(colour = layer_class,
                 group = layer_class),
             position = position_dodge(width = 0.25))+
  ggtitle("Salt Parameters")+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())


```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
salts<-lci %>% filter(category=="Metals",
                      RSLT_RESULT_SAMPLE_FRACTION=="TOTAL") %>% 
  mutate(layer_class=paste(info,class,sep="_"))

ggplot(salts,
       aes(x = simple,y = year)) + 
  geom_point(aes(colour = layer_class,
                 group = layer_class),
             position = position_dodge(width = 0.25))+
  ggtitle("Total Metals Parameters")+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())
```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
salts<-lci %>% filter(category=="Metals",
                      RSLT_RESULT_SAMPLE_FRACTION=="DISSOLVED") %>% 
  mutate(layer_class=paste(info,class,sep="_"))

ggplot(salts,
       aes(x = simple,y = year)) + 
  geom_point(aes(colour = layer_class,
                 group = layer_class),
             position = position_dodge(width = 0.25))+
  ggtitle("Dissolved Metals Parameters")+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())
```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
salts<-lci %>% filter(category=="Minerals") %>% 
  mutate(layer_class=paste(info,class,sep="_"))

ggplot(salts,
       aes(x = simple,y = year)) + 
  geom_point(aes(colour = layer_class,
                 group = layer_class),
             position = position_dodge(width = 0.25))+
  ggtitle("Minerals Parameters")+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())
```


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
salts<-lci %>% filter(category=="Trophic_Parameters") %>% 
  mutate(layer_class=paste(info,class,sep="_"))

ggplot(salts,
       aes(x = simple,y = year)) + 
  geom_point(aes(colour = layer_class,
                 group = layer_class),
             position = position_dodge(width = 0.25))+
  ggtitle("Trophic Parameters")+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())
```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
salts<-lci %>% filter(category=="Other_Nutrient_Parameters") %>% 
  mutate(layer_class=paste(info,class,sep="_"))

ggplot(salts,
       aes(x = simple,y = year)) + 
  geom_point(aes(colour = layer_class,
                 group = layer_class),
             position = position_dodge(width = 0.25))+
  ggtitle("Other Nutrient Parameters")+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
salts<-lci %>% filter(category=="In_Situ_Parameters") %>% 
  mutate(layer_class=paste(info,class,sep="_"))

ggplot(salts,
       aes(x = simple,y = year)) + 
  geom_point(aes(colour = layer_class,
                 group = layer_class),
             position = position_dodge(width = 0.25))+
  ggtitle("In Situ Parameters")+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())
```


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
salts<-lci %>% filter(category=="HABs_Parameters") %>% 
  mutate(layer_class=paste(info,class,sep="_"))

ggplot(salts,
       aes(x = simple,y = year)) + 
  geom_point(aes(colour = layer_class,
                 group = layer_class),
             position = position_dodge(width = 0.25))+
  ggtitle("HABs Parameters")+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```

## Lab Costs/Sample

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
costs<-read.csv("reconsidering.parameters.csv",na.strings=c("","NA"))
costs<-costs %>% 
  rename(param=CHARACTERISTIC_NAME) %>% 
  select(category,RSLT_RESULT_SAMPLE_FRACTION,simple,wqs_use,cost,DP,OW,BS,A,B) %>% distinct()
proposed1<-lci %>% filter(year=="2021",category!="HABs_Parameters") %>% 
  select(category,simple,RSLT_RESULT_SAMPLE_FRACTION) %>% distinct()
proposed2<-lci %>% filter(year=="2022",category=="HABs_Parameters") %>% 
  select(category,simple,RSLT_RESULT_SAMPLE_FRACTION) %>% distinct()
proposed<-merge(proposed1,proposed2,all=TRUE)
costs<-merge(proposed,costs,by=c('category','simple','RSLT_RESULT_SAMPLE_FRACTION'),all.x=TRUE)

costs<-costs %>% 
  mutate(DP=ifelse(DP==1,cost,DP),
         OW=ifelse(OW==1,cost,OW),
         BS=ifelse(BS==1,cost,BS),
         DP=ifelse(is.na(DP),0,DP),
         OW=ifelse(is.na(OW),0,OW),
         BS=ifelse(is.na(BS),0,BS),
         A=ifelse(is.na(A),0,A),
         B=ifelse(is.na(B),0,B),
         A_both=ifelse(A==1,DP+OW+BS,0),
         B_both=ifelse(B==1,DP+OW+BS,0),
         A_epi=ifelse(A==1,DP+OW,0),
         B_epi=ifelse(B==1,DP+OW,0)) %>% 
  select(category,simple,RSLT_RESULT_SAMPLE_FRACTION,wqs_use,A_both,A_epi,B_both,B_epi) %>% distinct()
wqs_costs21<-costs %>% 
  group_by(wqs_use) %>% 
  summarize(add_Ab=sum(A_both),
            add_Ae=sum(A_epi),
            add_Bb=sum(B_both),
            add_Be=sum(B_epi)) %>% 
  ungroup() %>% 
  filter(!is.na(wqs_use))

category_costs21<-costs %>% 
  group_by(category) %>% 
  summarize(add_Ab=sum(A_both),
            add_Ae=sum(A_epi),
            add_Bb=sum(B_both),
            add_Be=sum(B_epi)) %>% 
  ungroup()%>% 
  filter(!is.na(category))
```


Cost breakdown for 2022 proposed sampling parameters. 
Values in () are the increase from 2021.

Organized by Designated Use Categories. The () indicate the increase from 2021.

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
costs<-read.csv("reconsidering.parameters.csv",na.strings=c("","NA"))
costs<-costs %>% 
  rename(param=CHARACTERISTIC_NAME) %>% 
  select(category,RSLT_RESULT_SAMPLE_FRACTION,simple,wqs_use,cost,DP,OW,BS,A,B) %>% distinct()
proposed<-lci %>% filter(year=="2022") %>% 
  select(category,simple,RSLT_RESULT_SAMPLE_FRACTION) %>% distinct()
costs<-merge(proposed,costs,by=c('category','simple','RSLT_RESULT_SAMPLE_FRACTION'),all.x=TRUE)

costs<-costs %>% 
  mutate(DP=ifelse(DP==1,cost,DP),
         OW=ifelse(OW==1,cost,OW),
         BS=ifelse(BS==1,cost,BS),
         DP=ifelse(is.na(DP),0,DP),
         OW=ifelse(is.na(OW),0,OW),
         BS=ifelse(is.na(BS),0,BS),
         A=ifelse(is.na(A),0,A),
         B=ifelse(is.na(B),0,B),
         A_both=ifelse(A==1,DP+OW+BS,0),
         B_both=ifelse(B==1,DP+OW+BS,0),
         A_epi=ifelse(A==1,DP+OW,0),
         B_epi=ifelse(B==1,DP+OW,0)) %>% 
  select(category,simple,RSLT_RESULT_SAMPLE_FRACTION,wqs_use,A_both,A_epi,B_both,B_epi) %>% distinct()
wqs_costs<-costs %>% 
  group_by(wqs_use) %>% 
  summarize(A_both=sum(A_both),
            A_epi=sum(A_epi),
            B_both=sum(B_both),
            B_epi=sum(B_epi)) %>% 
  ungroup() %>% 
  filter(!is.na(wqs_use))
#add in 2021 costs for comparison
wqs_costs<-merge(wqs_costs,wqs_costs21,by=c('wqs_use'),all=TRUE)
wqs_costs<-wqs_costs %>% 
  replace_na(list(add_Ab=0,add_Ae=0,add_Bb=0,add_Be=0)) %>% 
  mutate(add_Ab=A_both-add_Ab,
         add_Ae=A_epi-add_Ae,
         add_Bb=B_both-add_Bb,
         add_Be=B_epi-add_Be)  %>% 
  bind_rows(summarise_all(., ~if(is.numeric(.)) sum(.) else "Total")) %>% 
  mutate(A_both=ifelse(add_Ab!=0,paste(A_both,"(",add_Ab,")",sep=""),A_both),
         A_epi=ifelse(add_Ae!=0,paste(A_epi,"(",add_Ae,")",sep=""),A_epi),
         B_both=ifelse(add_Bb!=0,paste(B_both,"(",add_Bb,")",sep=""),B_both),
         B_epi=ifelse(add_Be!=0,paste(B_epi,"(",add_Be,")",sep=""),B_epi)) %>% 
  select(wqs_use,A_both,A_epi,B_both,B_epi) %>% distinct()

DT::datatable(wqs_costs, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
 
```


Organized by Parameter Category. The value in the () indicate the increase from 2021.
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
category_costs<-costs %>% 
  group_by(category) %>% 
  summarize(A_both=sum(A_both),
            A_epi=sum(A_epi),
            B_both=sum(B_both),
            B_epi=sum(B_epi)) %>% 
  ungroup()

#add in 2021 costs for comparison
category_costs<-merge(category_costs,category_costs21,by=c('category'),all=TRUE)
category_costs<-category_costs %>% 
  replace_na(list(add_Ab=0,add_Ae=0,add_Bb=0,add_Be=0)) %>% 
  mutate(add_Ab=A_both-add_Ab,
         add_Ae=A_epi-add_Ae,
         add_Bb=B_both-add_Bb,
         add_Be=B_epi-add_Be)  %>% 
  bind_rows(summarise_all(., ~if(is.numeric(.)) sum(.) else "Total")) %>% 
  mutate(A_both=ifelse(add_Ab!=0,paste(A_both,"(",add_Ab,")",sep=""),A_both),
         A_epi=ifelse(add_Ae!=0,paste(A_epi,"(",add_Ae,")",sep=""),A_epi),
         B_both=ifelse(add_Bb!=0,paste(B_both,"(",add_Bb,")",sep=""),B_both),
         B_epi=ifelse(add_Be!=0,paste(B_epi,"(",add_Be,")",sep=""),B_epi)) %>% 
  select(category,A_both,A_epi,B_both,B_epi) %>% distinct()


DT::datatable(category_costs, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
```

## Lab Costs - TOTAL

The following is a budget of lab costs for processing the 2022 samples.
PLEASE NOTE: there are 6 sites in the probability based sampling list that are tentatively being sampled by CSLAP this year. We are still intending to sample them for LCI because CSLAP's parameter list is not the same. 

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
rm(list=setdiff(ls(), c("newdata","category_costs")))
sites<-read.csv("LCI.Field.Season.sites.csv",na.strings=c("","NA"))
sites<-sites %>% 
  filter(!is.na(PURPOSE),
         PURPOSE!="overdraw_Probability Sampling 2021-2025") %>% 
  select(PURPOSE,Project,LOCATION_ID,in.CSLAP,X_Coordinate,Y_Coordinate) %>% distinct() %>% 
  #repeat the one NLA site
  add_row(PURPOSE="NLA",Project="NLA3",LOCATION_ID="0904LON0301_DH",in.CSLAP=NA,
          X_Coordinate=-75.06178,Y_Coordinate=44.27153) %>% 
  mutate(Project=ifelse(Project=="NLA1 and NLA3 (MUST REPEAT)","NLA1",Project),
         #create a centroid location to get classification next
         LOCATION_HISTORY_ID=gsub("_DH","_C",LOCATION_ID))

#pull class
class<-newdata %>% select(LOCATION_HISTORY_ID,LOCATION_WATERBODY_CLASSIFICATION) %>% distinct()
sites<-merge(sites,class,by=c('LOCATION_HISTORY_ID'),all.x=TRUE)
#add costs
sites<-sites %>% 
  rename(class=LOCATION_WATERBODY_CLASSIFICATION) %>% 
  mutate(cost=ifelse(grepl("A",class),632,546),
         cost=ifelse(PURPOSE=="NLA",0,cost),
         cost=ifelse(PURPOSE=="Probability Sampling 2021-2025",cost+325,cost)) 

sites<-sites %>% 
  group_by(PURPOSE) %>% 
  summarize(Number.of.Lakes=n(),
            Number.of.Events=n(),
            sample_cost=sum(cost)) %>% 
  ungroup() 


QA<-read.csv("staff.allocations.csv",na.strings=c("","NA"))
QA<-QA %>% 
  rename(days=X..days,
         staff=X..staff.per.team) %>% 
  filter(!is.na(PURPOSE)) %>% 
  select(PURPOSE,tripname,lodging,MI.E,Month,days,teams,staff) %>% distinct() %>% 
  group_by(PURPOSE) %>% 
  summarize(Num.QA=sum(teams)) %>% 
  ungroup() %>% 
  mutate(QA_cost=Num.QA*632) 
sites<-merge(sites,QA,by=c('PURPOSE'),all=TRUE)

sites<-sites %>%  
  mutate(cooler_cost=ifelse(PURPOSE=="Probability Sampling 2021-2025",(Number.of.Events/2)*15,0),
         cooler_cost=ifelse(PURPOSE=="PWS Screening",(Number.of.Events/3)*28,cooler_cost)) %>% 
  mutate(Total=sample_cost+QA_cost+cooler_cost)%>% 
  bind_rows(summarise_all(., ~if(is.numeric(.)) sum(.) else "Total"))

 DT::datatable(sites, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
lab<-sites

```

## Travel Costs

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
rm(list=setdiff(ls(), c("newdata","lab","category_costs")))
sites<-read.csv("staff.allocations.csv",na.strings=c("","NA"))
sites<-sites %>% 
  rename(days=X..days,
         staff=X..staff.per.team) %>% 
  filter(!is.na(PURPOSE)) %>% 
  select(PURPOSE,tripname,lodging,MI.E,Month,days,teams,staff) %>% distinct() %>% 
  mutate(teams=ifelse(staff==4,teams*2,teams),
         cost=(((2*lodging)+(3*MI.E))*teams*2),
         cost=ifelse(days==2,((1*lodging)+(2*MI.E))*teams*2,cost)) %>% 
  group_by(PURPOSE) %>% 
  summarize(cost=sum(cost)) %>% 
  ungroup()%>% 
  bind_rows(summarise_all(., ~if(is.numeric(.)) sum(.) else "Total"))

DT::datatable(sites, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
travel<-sites
```

## Totals

It's worth noting that costs are less than predicted. 
This is in part because we are sampling far fewer events than in 2021 (181 events in 2021 compared to the planned 102 events in 2022). 
Also, we aren't paying for the NLA lab costs.
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

travel1<-travel %>% filter(PURPOSE=="Total") %>% distinct() %>% select(PURPOSE,cost)
travel1<-travel1 %>% rename(Travel=cost)
lab1<-lab %>% filter(PURPOSE=="Total") %>% distinct() %>% select(PURPOSE,Total) %>% distinct()
lab1<-lab1 %>% rename(Lab=Total)
total<-merge(travel1,lab1,by=c('PURPOSE'))
total<-total %>% 
  rename(Costs=PURPOSE) %>% 
  mutate(Supplies=6000,
         Equipment=20000,
         Total=Lab+Travel+Supplies+Equipment) %>% 
  gather(Category,cost) %>% 
  filter(Category!="Costs") 

DT::datatable(total, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

```